Deep Reinforcement Learning for Conservation Decisions

### 2022-02-03 manuscript/manuscript.Rmd
Looking like the final version here

---
title: Deep Reinforcement Learning for Conservation Decisions
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mlapeyro@berkeley.edu
  - name: Melissa S. Chapman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mchapman@berkeley.edu
  - name: Kari E. A. Norman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: kari.norman@berkeley.edu
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu (corresponding author)
abstract: |
  Can machine learning help us make better decisions about a changing planet?
  In this paper, we illustrate and discuss the potential of a promising corner of machine learning known as _reinforcement learning_ (RL) to help tackle the most challenging conservation decision problems.
  RL is uniquely well suited to conservation and global change challenges for three reasons: 
  (1) RL explicitly focuses on designing an agent who _interacts_ with an environment which is dynamic and uncertain,
  (2) RL approaches do not require massive amounts of data, 
  (3) RL approaches would utilize rather than replace existing models, simulations, and the knowledge they contain.
  While this potential may be compelling, the challenges involved in successfully deploying RL-based management to realistic scenarios are formidable --
  the required expertise and computational cost may place these applications beyond the reach of all but large, international technology firms.
  Ecologists must establish a better understanding of how these algorithms work and fail if we are to realize this potential and avoid the pitfalls such a transition would bring.
  To this end, we provide a conceptual and technical introduction to RL and its relevance to ecological and conservation challenges, including examples of a problem in setting fisheries quotas and in managing ecological tipping points.
  Four appendices with annotated code provide a tangible introduction to researchers looking to adopt, evaluate, or extend these approaches.

keywords:
  - Reinforcement Learning
  - Conservation
  - Machine Learning
  - Artificial Intelligence
  - Tipping points
bibliography: references.bib
nocite: |
    @A2C, @trpo, @ddpg, @TD3, @impala, @hafner_learning_2019, @weber_imagination-augmented_2018, @kirk_survey_2022
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{lineno}
  - \usepackage{setspace}
#  - \usepackage{endfloat}
#  - \linenumbers
#  - \doublespacing

@@

the number of figures, tables, and text boxes: **6** (five figures, 1 table)

@@

We do recognize the need for more complex, "precise" models to derive quantitatively precise management policy in specific management problems such as fisheries, forestry, or climate change mitigation efforts. 
The examples here should provide ecologists with a greater background to engage in the collaborative development of such methods, while also highlighting challenges which ecological problems pose to existing techniques.
Even the simple models we consider here contain sources of complexity such as hysteresis that are likely commonplace to ecological problems but have not been well represented in RL research to date. 
Applying RL to a range of "precise" models is currently beyond the scope of any one paper: it will require not only more computational resources, but more collaborations between large teams of ecologists and computer scientists.

# RL overview

@@

>>>>

---
title: Deep Reinforcement Learning for Conservation Decisions
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mlapeyro@berkeley.edu
  - name: Melissa S. Chapman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: mchapman@berkeley.edu
  - name: Kari E. A. Norman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: kari.norman@berkeley.edu
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu (corresponding author)
abstract: |
  1. Can machine learning help us make better decisions about a changing planet? In this paper, we illustrate and discuss the potential of a promising corner of machine learning known as deep reinforcement learning (RL) to help tackle the most challenging conservation decision problems. We provide a conceptual and technical introduction to deep RL as well as annotated code so that researchers can adopt, evaluate and extend these approaches.
  2. RL explicitly focuses on designing an agent who interacts with an environment which is dynamic and uncertain. Deep RL is the subfield of RL that incorporates deep neural networks into the agent. We train deep RL agents to solve sequential decision-making problems in setting fisheries quotas and managing ecological tipping points.
  3. We show that a deep RL agent is able to learn a nearly optimal solution for the fisheries management problem. For the tipping point problem, we show that a deep RL agent can outperform a sensible rule-of-thumb strategy. 
  4. Our results demonstrate that deep RL has the potential to solve challenging decision problems in conservation. While this potential may be compelling, the challenges involved in successfully deploying RL-based management to realistic scenarios are formidable -- the required expertise and computational cost may place these applications beyond the reach of all but large, international technology firms. Ecologists must establish a better understanding of how these algorithms work and fail if we are to realize this potential and avoid the pitfalls such a transition would bring. We ultimately set forth a research framework based on well-posed, public challenges so that ecologists and computer scientists can collaborate towards solving hard decision-making problems in conservation.

keywords:
  - Artificial Intelligence
  - Conservation
  - Machine Learning
  - Reinforcement Learning
  - Tipping points
bibliography: references.bib
nocite: |
    @A2C, @trpo, @ddpg, @TD3, @impala, @hafner_learning_2019, @weber_imagination-augmented_2018, @kirk_survey_2022
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{lineno}
  - \usepackage{setspace}
  - \linenumbers
  - \doublespacing

@@

the number of figures, tables, and text boxes: **7** (5 figures, 2 tables)

@@

We do recognize the need for more complex, "precise" models to derive quantitatively precise management policy in specific management problems such as fisheries, forestry, or climate change mitigation efforts. 
The examples here should provide ecologists with a greater background to engage in the collaborative development of such methods, while also highlighting challenges which ecological problems pose to existing techniques.
Even the simple models we consider here contain sources of complexity such as hysteresis that are likely commonplace to ecological problems but have not been well represented in RL research to date. 
Applying RL to a range of "precise" models is currently beyond the scope of any one paper: it will require not only more computational resources but more collaborations between large teams of ecologists and computer scientists.

# Materials and Methods

@@

# Acknowledgements {-}

All authors were supported by the NSF Faculty Early Career Development Program Award 1942280. We thank Benjamin Blonder, Zachary Sunberg, and Claire Tomlin for their thoughtful comments.

# Conflicts of Interest {-}

The authors declare there are no conflicts of interest.

# Authors' Contributions {-}

M.L. and C.B. developed the code and wrote the manuscript with support from K.E.A.N. and M.S.C.

# Data Availability {-}

We have created a github repository <https://github.com/boettiger-lab/rl-intro> that contains the code used to produce the figures herein.
